Relationships in the extremes and their influence on competition and coexistence

### 2021-06-29 paper/Paper.Rnw
made additions to sup mat to support Methods - still working on it though

%Put other possible titles below, commented, and we will select the best one at the end
%Relationships between extreme events and their influence on competition and coexistence

\author[a]{Jasmin Albert}
\author[a,b]{Daniel C. Reuman}
%***DAN: Jasmin, you should choose one way your name will appear, exactly, on all your publications,
%so everyone knows its you. For instance, I always appear as Daniel C. Reuman, not Daniel Reuman, not
%Dan Reuman, etc. This makes it possible to find all my papers by searching. So please decide what
%name to use, Jasmin Albert or Pimsupa Albert or P. Jasmin Albert or whatever. Also keep in mind
%that because "Albert" is a common name you might want something less common for the rest of it 
%so that your papers are not confused with anyone elses, though of course I emphasize that how you
%are going to be known it totally up to you, I'm just recommending you think it through now so you
%can be consistent. I recognize that this all may be moot if you one day change your name and want to
%be known by the new name on papers after that point, but anyway in case you don't ever change your 
%name you might as well be consistent now, at least. So please change your name above to whatever
%you want to be officially known as on scientific papers. Same change in the sup mat, please.

@@

Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
in the last para of the Intro, so please incorporate these into your last para. If you have a better way
to phrase these, please put both so we can compare at some point when we meet.
G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
from ATA compared to other mechanisms.

@@

In this section we pursue goal G1 from the Introduction: to augment the MTC to show how ATA
plays a role in species coexistence.
%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
For pedgogical clarity, we develop each step of the theory in general and for a simple classical model 
(the lottery model), as an example.

The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundiity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

@@

For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/\delta N$, which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.

@@

For the lottery model, when $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SI-SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
Since we assumed log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%***DAN: This statement also needs adjusting if you go with q_ij=1
So $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).

We now specify how to decompose storage effects into
contributions due to asymmetric tail associations and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between invader and resident mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components.
The term $\Delta_i^{[EC]}$ is the component due to $EC$ correlation \emph{per se}, after 
removing the effects of asymmetric tail associations. 
The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
resident growth rates (see below).

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume that the cummulative distribution functions (cdfs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the cdf of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\tilde{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\tilde{\rho}_i$ is then determined 
in one of 
four alternative ways: $\tilde{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\tilde{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\tilde{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\tilde{b}_i),\exp(\tilde{b}_j))$, where $(\tilde{b}_i,\tilde{b}_j)$
is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\tilde{b}_i)=\var(\tilde{b}_j)=\sigma$,
and $\cov(\tilde{b}_i,\tilde{b}_j)=\cov(b_i,b_j)$. So $(\tilde{b}_i,\tilde{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
so $E_j^\nat=E_j$ and $C_{j \bs i}^\nat = C_{j \bs i}$. Thus $\olr_{j \bs i}=\olrn_{j \bs i}$.

The other mechanisms in modern coexistence theory that apply in our modelling context (of purely temporal variation)
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Mathematical details are given in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
coexistence opperating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution of environmental variation;
$\Delta_i^C$ is a contribution of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution of $EC$ covariation -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA is storage effects. Pairing our decomposition 
(\ref{eq:storage_effects_further1})-(\ref{eq:storage_effects_further}) with the 
previous MTC decomposition (\ref{eq:previous_decomp}) gives the extension of the MTC which was goal G1 of the Introduction.

@@

%Goals, for reference:
%G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
%plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
%We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
%models exhibiting solely temporal variation; spatial variation can be considered in future work.
%G2) To explore using a simple model the relative magnitude of the contributions of coexistence that come
%from ATA compared to other mechanisms.

@@

@@

>>>>

%Put other possible titles below, commented, and we will select the best one at the end
%Relationships between extreme events and their influence on competition and coexistence
%Coexistence and consequences of extreme environmental relationships
%Coexistence and consequences of environmental relationships in their extremes
%Coexistence and implications of asymmetric environmental relationships
%Extreme [climatic] events and their implications for biodiversity
%Deductions to coexistence, extreme events, and asymmetric relationships
%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence

\author[a]{Pimsupa Jasmin Albert}
\author[a,b]{Daniel C. Reuman}

@@

%Jasmin, I think the below makes a pretty good start at tryinbg to nail down a good way to express our goals
%in the last para of the Intro, so please incorporate these into your last para. If you have a better way
%to phrase these, please put both so we can compare at some point when we meet.
G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
from ATA compared to other mechanisms.

@@

We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
plays a role.
%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.

The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

@@

For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.

@@

For the lottery model, when $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SI-SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
%***DAN: Note this might be wrong. Some of the language in Ellner et al (2016) indicates q_{ij} here should be 1,
%even if \mu_1 and \mu_2 are distinct, as long as we use the same \delta for both species (which we do).
%But when I try to follow the definitions of the same paper to compute q_{ij} I get the quantity listed above,
%which is not 1. I'll need to get to the bottom of this, possibly by speaking to Ellner, and then put the right 
%value here, and also write the corresponding sup mat section accordingly (though you may not need that section, 
%it may be enough to just cite Ellner and Chesson if you go with q_ij=1.
Since we assumed log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
%***DAN: This statement also needs adjusting if you go with q_ij=1
Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).

We now specify how to decompose storage effects into
contributions due to asymmetric tail associations and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between invader and resident mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
asymmetries of tail association between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATA).
The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
having removed the effects of asymmetric tail associations.
%The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
%this is a decompostion of storage effects, which were denoted by \cite{Ellner_2019} using a superscript $(EC)$,
%i.e., $\Delta_i^{(EC)}$, to indicate that storage effects are, in turn, part of a decomposition of a difference between invader and
%resident growth rates (see below).  
The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decompostion within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term within the decomposition provided by the MTC \citep{Ellner_2019} -- see below for details.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume, for simplicity, that the cummulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
in one of 
four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further1}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.

The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{i \bs i}$ which is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
coexistence opperating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{i \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of the MTC which was goal G1 of the Introduction.

@@

%Goals, for reference:
%G1) To augment the modern theory of coexistence (MTC) to show how asymmetric tail association (ATA)
%plays a role. The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
%We address how ATA contributes to these components. We here examine mechanisms of coexistence that apply in 
%models exhibiting solely temporal variation; spatial variation can be considered in future work.
%G2) To explore using a simple model the relative magnitudes of the contributions to coexistence that come
%from ATA compared to other mechanisms.

@@

Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$,
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}); so we only plotted 
against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}); so we only generated plots 
for left-tail associated noise. For each combination of $\delta=W, X, Y, Z$
%***DAN: Jasmin, please select 4 sensible values, including \delta=1, and put the ones you select here
and $\mu_1-\mu_2=A,B,C,D$,
%***DAN: Jasmin, please select 3-4 sensible values and put the ones you select here
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model.
%***DAN: Jasmin, if you select 4 values for mu1-mu2 and 4 for delta, then Fig. 3 is going to have 16 panels arranged in a 
%4 by 4 grid, each of which will plot all the components against sigma. You'll need to use different colors and/or line
%types for the components and have a legend.
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section X.
%***DAN: Dan to add the reference to the right sup mat section once things get sufficiently organized in sup mat to know what section.

@@

\begin{figure}
\caption{Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick
model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should
discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader
is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through
the Theory section, which is inevitably going to be tough going for many readers.\label{fig:pedagog}}
\end{figure}
